{\normalsize \textbf{Vagueness and imprecise imitation in signalling games}}

### 2016-06-07 paper/version_final/paper.tex
moved floats to end (really?)
Small changes after final reading. Fixed some bibliography issues.

@@

Many concepts and expressions are vague. A vague category knows clear cases that fall under it,
clear cases that do not, and also so-called borderline cases. Borderline cases do not clearly
apply, nor do they clearly not apply, and there may be differences between borderline cases in
terms of how well they represent the category in question. Vagueness does not seem to
dramatically affect the success of everyday communication, but it is troublesome for some of
the most prominent theories of language and meaning. This is especially so for the
logico-positivist tradition of Frege, Russell, and early Wittgenstein, which is challenged by
the paradoxes vagueness gives rise to.

But there are other intriguing aspects about vagueness. One perplexing issue is how vagueness
could arise and be maintained in the first place. This is an apparent puzzle for functionalist
accounts that maintain that concepts and linguistic meanings evolved driven towards
efficiency. \citetbjps{Lipman2009:Why-is-Language} argues that, in common interest signalling
situations, the existence of unclear borderline cases entails inefficiency of categorisation or
communication, or at least no advantage. The challenge is then to explain how vagueness
can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii)
without undermining the possibility of evolving, learning and communicating with a meaningful
language. A number of authors have consequently tried to explain why vagueness evolved as
something that is itself useful
(e.g.~\citealtbjps{Jaegherde-Jaegher2003:A-Game-Theoreti}; \citealtbjps{Deemter2009:Utility-and-Lan}; \citealtbjps{BlumeBoard2013:Intentional-Vag}).
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citepbjpspre{e.g.}{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
learning strategies \citepbjpspre{e.g.}{OConnor2013:The-Evolution-o}. This paper contributes to the
latter line of thought.

We believe that vagueness in language and thought may have many reasons, not just one. We focus
here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
idea is that vagueness is, at least in part, due to imprecision in the perception of similar
stimuli and imprecision in the realisation of similar responses. On this view, vagueness in
language may be seen as a necessary sub-optimality due to limitations in another domain of
cognition, first and foremost perception. It could then be speculated that the whole system,
perception and language together, may be an almost optimal adaptation in a larger frame
of reference, e.g., if we take into account the metabolic costs for increased perceptual
accuracy. We will not engage in such speculation here. Rather we will explore the consequences
of perceptual limitations on processes of meaning evolution in a suitable formal framework. In
other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
by appeal to the naturalness of vagueness, it needs to be spelled out how exactly confusability
of stimuli could come into play in a process of meaning evolution and how it could lead to
vague but by-and-large informative signal meaning. The formal model that this paper introduces
does exactly that. But it also does more. We find that confusability of stimuli can regularise
and systematise evolving meaning. This suggests a possibly advantageous side-effect that a
natural cause of vagueness may have and that might compensate some of the disadvantages that
vagueness may have for communicative efficiency \citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.

@@

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalisation of the replicator
dynamic which is derived from the idea that agents imitate other agents' behaviour while
possibly confusing similar states.\footnote{Much previous work has investigated the interplay
  of adaptive dynamics, including imitation-based update protocols, on the one hand, and noise
  or mutation, on the other hand (e.g., \citealtbjps{FosterYoung1990:Stochastic-Evol};
  \citealtbjps{FudenbergHarris1992:Evolutionary-Dy};
  \citealtbjps{KandoriMailath1993:Learning-Mutati}; \citealtbjps{Young1993:The-Evolution-o};
  \citealtbjps{FudenbergImhof2006:Imitation-Proce}). While this line of research often looks at
  finite populations and the extreme long-term behavior of the system under generic randomness,
  the focus here is on a quite particular source of stochastic noise and its quite particular
  role in the evolution of meanings through signalling.}  This imprecise imitation dynamic is
explored in Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares
our approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.

@@

Signalling games, as introduced by \citetbjps{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signalling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behaviour can evolve is a highly interesting
topic that we are only beginning to fully understand
(e.g.~\citealtbjps{Warneryd1993:Cheap-Talk-Coor}; \citealtbjps{BlumeKim1993:Evolutionary-St}; \citealtbjps{Huttegger2007:Evolution-and-t}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{Barrett2009:The-Evolution-o}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}; \citealtbjps{Skyrms2010:Signals}).

@@

There are at least two potential ways of interpreting the signalling setup.  Sender and
receiver can be distinct entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why linguistic categories are
well-behaved and orderly in the way they appear to be. More abstractly, sender and receiver can
also be thought of as distinct modules in a single system, where the first module must
discretise the information it is fed by selecting a small sample of, suggestively, category
labels. These are passed to a second module that tries to decode the original information. In
this case, evolving Voronoi languages would explain why conceptual categories are well-behaved
and orderly in the way that they appear to be.  Seen in this light, sim-max games may provide a
foundation to approaches in cognitive semantics that rely on the notion of conceptual spaces.
\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued
that natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorisation system can be obtained: storing the prototypes lets us recover the categories
without having to store each category's extension. However, what is left unexplained is where
the prototypes come from, and why we would not see just any distribution of prototypes as an
equally efficient classification system. This is where sim-max games can contribute a
principled approach to deriving, in an independent way, not only convex categories but also
prototypical exemplars belonging to them.  These ideas, and more, are developed further in
(\citealtbjps{Jager2007:The-Evolution-o};
\citealtbjps{JagerRooijvan-Rooij2007:Language-Struct};
\citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}), among
others.

\subsection{Vagueness in sim-max games and conceptual spaces}

This outline of an approach to categorisation using sim-max games
leaves many problems unaddressed. One of them is that natural
categories for continuously variable stimuli usually do not have
unique, point-valued prototypes and clear category boundaries. We
would like to account for the possibility of such vagueness, in
particular: (i) clear positive examples of a vague category should
show a gradient transition to clear negative examples;
(ii) prototypes should likewise be gradient regions, peaking at the
centre of the vague category they represent.

@@

Alternative approaches are taken by, for example, \citetbjps{FrankeJager2010:Vagueness-Signa} and
\citetbjps{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
met by evolving strategies in sim-max games. To illustrate what vague signalling would look
like, let us briefly consider a sim-max game with six equiprobable states and two messages and
what its equilibria would be like (see \citealtbjps{Jager2007:The-Evolution-o} and \citealtbjps{OConnor2013:The-Evolution-o} for a more thorough discussion of equilibria of
sim-max games). We assume that utilities
are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
three pairs of sender and receiver strategies. States are arranged according to their
similarity: the closer they are to each other, the more similar they are. The pair in
Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
signals is sub-optimal given the receiver's behaviour. Namely, if \mymessg{1} is interpreted as
\mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
sending \mymessg{1} in \mystate{3} than from sending \mymessg{2}, because (by assumption)
\mystate{3} is more similar to \mystate{2} than \mystate{5} is. In contrast,
Figure~\ref{fig:efficientEq} shows a maximally efficient equilibrium. This is a partial pooling
equilibrium in the sense that the speaker uses the same message for several states. Partial
pooling equilibria can be less inefficient than other non-pooling equilibria, if there are
enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
the two-message case. So, while partial pooling may entail inefficiency in some sense, and
while partial pooling can hamper the evolution of maximally efficient signal use
(e.g.~\citealtbjps{Hutteger:2007-Evol-Indicatives-Imperatives}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}),
this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets
close, so to speak. It shows smooth transitions across similar states at the boundaries of
categories and across acts around the most prototypical instances (as indicated by decreasing
thickness of arrows in Figure~\ref{fig:vagueUse}).

@@

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
an example for the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, `accumulate' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of transmission biases are isolated, as in iterated learning models
(\citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
or when they interact with functional pressure towards efficient communication, e.g., as
formalised in the replicator mutator dynamic
(e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}).

The emphasis of previous models that studied the effects of transmission infidelity on the
evolution of language has been on inductive biases and the systematicity, compression and
regularisation that they can introduce. Here, we would like to show that transmission noise of
a different kind can lead to regularisation as well and also give rise to vague meaning. We
find that shared perceptual biases that perturb the transmission of successful signalling
behaviour can regularise, facilitate and accelerate the evolution of meaning conventions, albeit
at the cost of vagueness. Concretely, we formalise the expected change in the behaviour of a
population of agents that try to imitate other agents' signalling behaviour. We assume, however,
that both realisation of behaviour, as well as observability of others' behaviour, is
systematically perturbed by noise. The resulting population-level dynamic generalises the
replicator dynamic \citepbjps{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
evolutionary dynamic based on imitation
(e.g.~\citealtbjps{Helbing1996:A-Stochastic-Be}; \citealtbjps{Schlag1998:Why-Imitate-and}). The inclusion of
confusability of stimuli does not undermine the possibility of evolving communicative signalling
behaviour. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
emergence of communicative signalling because it unifies and regularises evolutionary outcomes,
making it appear as if agents were applying inductive biases or generalising over partial
observations, when in truth this is the sole effect of confusion of similar stimuli.

@@

The expected utility of choices at each choice point is:\footnote{The notation
  $\propto$, for `proportional to', that is used here and hereafter means that the right-hand
  side might still need to be normalised so as to have probabilities sum to one. In general,
  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
    \States} P(\state \probbar \messg)  \myts
  U(\state, \state')\,, \text{\ \  where } &
P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
sum, for each interpretation $\state'$, of the probability that some random receiver will
choose $\state'$ given $\messg$ times the actual utility of pair $\state$ and
$\state'$.

The discrete-time replicator dynamic tracks changes in frequency of choices in the population
as proportional to their expected utilities
\citepbjpspre{e.g.}{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
  utilities are always positive.}
\begin{align}
  \label{eq:RDDiscrete}
  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto \Rstrat(\state \probbar \messg) \myts
    \EU(\state, \messg,\Sstrat)\,.   
\end{align}
For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
average agent in the population after the update is proportional to the probability of
seeing it before the update, which is $\Sstrat(\messg \probbar \state)$, times the
expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
by a gradient of current frequency and a measure of how good they are.

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioural dispositions in a population of signallers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioural adaptation. We focus
here on one of the simplest: imitation of success
\citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following (see
Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
update scheme).  Every now and then a random agent gets a chance to alter his behaviour for one
of his choice points. Say, this is a sender who gets to `reconsider' his choice of message
for state $\state$. The revising agent then observes what some random agent does at $\state$,
say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
expected utility of $\messg$ for state $\state$.

@@

Imitation of success, as described above, presupposes that agents make no mistakes when
observing states, or choosing interpretations. This may not always be an appropriate
assumption, especially when some states can be perceptually similar and therefore likely to be
confused for one another.
Human performance in these situations has been studied by a number of authors in experimental psychology. %~\citepbjps{takane_structures_1992}.
In a stimulus identification experiment~\citepbjps{luce-detection-1963}, subjects are presented in each trial with a stimulus to be identified out of a fixed set.
The more similar the stimuli are to each other, the larger the number of errors subjects make.
For example, in Robert Nosofsky's experiments~\citepbjps{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the centre to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
In over 9000 trials, the two subjects identified the stimulus correctly only approximately 44\% and 35\% of the time.
Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colours, and magnitude of lines and areas~\citepbjps{donkin-why-2015}.
The empirical data confirms not only the pervasiveness of variation in the subjects' ability to correctly identify stimuli, but also the relation between similarity and likelihood of stimulus confusion, and further suggests that the phenomenon might extend to all types of perception.

@@

The presence of observation and realisation noise also affects imitation of successes. Here, we
focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:imit-succ-with}. Take a
sender who gets to revise behaviour for state $\state$. What agents can plausibly revise by
imitation is their pure strategy, which maps perceived states onto messages. When an agent gets
to revise his strategy for the perceived state $\state$, this need not necessarily be the
actual state. Moreover, when that agent observes $\state$, another agent may perceive yet a
different state. Given what that latter agent perceives, his actual (pure) strategy will
determine what he plays. In sum, to describe how likely the potential imitator observes a
message choice $\messg$, we are interested in the conditional probability
$P_o(\messg \probbar \state)$ that some other random agent selects $\messg$ when the first
agent perceives $\state$. This $P_o(\messg \probbar \state)$ is derived from the prior
probability of states, the current sender population behaviour $\Sstrat$ and the given
observation noise (see Appendix~\ref{sec:imit-succ-with}). Eventually, the imitating agent adopts
$\messg$ as his choice for $\state$ with a probability given by the expected utility of sending
$\messg$ when perceiving state $\state$. Expected utility should, of course, take the
probabilistic confusability of states into account as well.

>>>>

\usepackage[nomarkers, nolists]{endfloat}

@@

Many concepts and expressions are vague. A vague category knows clear cases that fall under it,
clear cases that do not, and also so-called borderline cases. Borderline cases do not clearly
belong to the category, nor do they clearly not belong, and there may be differences between borderline cases in
terms of how well they represent the category in question. Vagueness does not seem to
dramatically affect the success of everyday communication, but it is troublesome for some of
the most prominent theories of language and meaning. This is especially so for the
logico-positivist tradition of Frege, Russell, and early Wittgenstein, which is challenged by
the paradoxes vagueness gives rise to.

There are other intriguing aspects about vagueness. One perplexing issue is how vagueness
could arise and be maintained in the first place. This is an apparent puzzle for functionalist
accounts that maintain that concepts and linguistic meanings evolved driven towards
efficiency. \citetbjps{Lipman2009:Why-is-Language} argues that, in common interest signalling
situations, the existence of unclear borderline cases entails inefficiency of categorisation or
communication, or at least no advantage. The challenge is then to explain how vagueness
can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii)
without undermining the possibility of evolving, learning and communicating with a meaningful
language. A number of authors have consequently tried to explain why vagueness evolved as
something that is itself useful
(e.g.~\citealtbjps{Jaegherde-Jaegher2003:A-Game-Theoreti}; \citealtbjps{Deemter2009:Utility-and-Lan}; \citealtbjps{BlumeBoard2013:Intentional-Vag}).
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citepbjpspre{e.g.}{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
learning strategies \citepbjpspre{e.g.}{OConnor2013:The-Evolution-o}. This paper contributes to the
latter line of thought.

We believe that vagueness in language and thought may have many reasons, not just one. We focus
here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
idea is that vagueness is, at least in part, due to imprecision in the perception of similar
stimuli and imprecision in the realisation of similar responses. On this view, vagueness in
language may be seen as a necessary sub-optimality due to limitations in another domain of
cognition, first and foremost perception. It could then be speculated that the whole system,
perception and language together, may be an almost optimal adaptation in a larger frame
of reference, for example if we take into account the metabolic costs for increased perceptual
accuracy. We will not engage in such speculation here. Rather we will explore the consequences
of perceptual limitations on processes of meaning evolution in a suitable formal framework. In
other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
by appeal to the naturalness of vagueness, one needs to spell out how exactly could confusability
of stimuli come into play in a process of meaning evolution and how it could lead to
vague but by-and-large informative signal meaning. The formal model that this paper introduces
does exactly that. But it also does more. We find that confusability of stimuli can regularise
and systematise evolving meaning. This suggests a possibly advantageous side-effect that a
natural cause of vagueness may have and that might compensate some of the disadvantages that
vagueness may have for communicative efficiency \citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.

@@

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalisation of the replicator
dynamic which is derived from the idea that agents imitate other agents' behaviour while
possibly confusing similar states.\footnote{Much previous work has investigated the interplay
  of adaptive dynamics, including imitation-based update protocols, on the one hand, and noise
  or mutation, on the other hand (e.g.~\citealtbjps{FosterYoung1990:Stochastic-Evol};
  \citealtbjps{FudenbergHarris1992:Evolutionary-Dy};
  \citealtbjps{KandoriMailath1993:Learning-Mutati}; \citealtbjps{Young1993:The-Evolution-o};
  \citealtbjps{FudenbergImhof2006:Imitation-Proce}). While this line of research often looks at
  finite populations and the extreme long-term behavior of the system under generic randomness,
  the focus here is on a quite particular source of stochastic noise and its quite particular
  role in the evolution of meanings through signalling.}  This imprecise imitation dynamic is
explored in Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares
our approach to that of others. Appendix~\ref{sec:formal-stuff} provides formal detail.

@@

Signalling games, as introduced by \citetbjps{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signalling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behaviour can evolve is a
topic that we are only beginning to fully understand
(e.g.~\citealtbjps{Warneryd1993:Cheap-Talk-Coor}; \citealtbjps{BlumeKim1993:Evolutionary-St}; \citealtbjps{Huttegger2007:Evolution-and-t}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{Barrett2009:The-Evolution-o}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}; \citealtbjps{Skyrms2010:Signals}).

@@

There are at least two potential ways of interpreting the signalling setup.  Sender and
receiver can be distinct entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why linguistic categories are
well-behaved and orderly in the way they appear to be. More abstractly, sender and receiver can
also be thought of as distinct modules in a single system, where the first module must
discretise the information it is fed by selecting a small sample of, suggestively, category
labels. These are passed to a second module that tries to decode the original information. In
this case, evolving Voronoi languages would explain why conceptual categories are well-behaved
and orderly in the way that they appear to be.  Seen in this light, sim-max games may provide a
foundation to approaches in cognitive semantics that rely on the notion of conceptual spaces.
\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued
that natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorisation system can be obtained: storing the prototypes lets us recover the categories
without having to store each category's extension. However, what is left unexplained is where
the prototypes come from, and why we would not see just any distribution of prototypes as an
equally efficient classification system. This is where sim-max games can contribute a
principled approach to deriving, in an independent way, not only convex categories but also
prototypical exemplars belonging to them.  These ideas, and more, are developed further by
\citetbjps{Jager2007:The-Evolution-o}, \citetbjps{JagerRooijvan-Rooij2007:Language-Struct}, \citetbjps{JagerMetzger2011:Voronoi-Languag}, \citetbjps{OConnor2014-OCOEPC}, among
others.

\subsection{Vagueness in sim-max games and conceptual spaces}
\label{sec:vagueness-in-sim-max-games}

This outline of an approach to categorisation using sim-max games
leaves some problems unaddressed. One of them is that, usually, natural
categories for continuously variable stimuli neither have
clear boundaries, nor do they have unique, point-valued prototypes. We
would like to account for the possibility of such vagueness, in
particular: (i) clear positive examples of a vague category should
show a gradient transition to clear negative examples;
(ii) prototypes should likewise be gradient regions, peaking at the
centre of the vague category they represent.

@@

Alternative approaches are taken by, for example, \citetbjps{FrankeJager2010:Vagueness-Signa} and
\citetbjps{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
met by evolving strategies in sim-max games. To illustrate what vague signalling would look
like, let us briefly consider a sim-max game with six equiprobable states and two messages and
what its equilibria would be like (see \citealtbjps{Jager2007:The-Evolution-o} and \citealtbjps{OConnor2013:The-Evolution-o} for a more thorough discussion of equilibria of
sim-max games). We assume that utilities
are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
three pairs of sender and receiver strategies. States are arranged according to their
similarity: the closer they are to each other, the more similar they are. The pair in
Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
signals is sub-optimal given the receiver's behaviour. Namely, if \mymessg{1} is interpreted as
\mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
sending \mymessg{1} in \mystate{3} than from sending \mymessg{2}, because (by assumption)
\mystate{3} is more similar to \mystate{2} than \mystate{5} is. In contrast,
Figure~\ref{fig:efficientEq} shows a maximally efficient equilibrium. This is a partial pooling
equilibrium in the sense that the speaker uses the same message for several states. Partial
pooling equilibria can be less inefficient than other non-pooling equilibria, if there are
enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
the two-message case. So, while partial pooling may entail inefficiency in some sense, and
while partial pooling can hamper the evolution of maximally efficient signal use
(\citealtbjps{Hutteger:2007-Evol-Indicatives-Imperatives}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}),
this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets
close, so to speak. It shows smooth transitions across similar states at the boundaries of
categories and across acts around the most prototypical instances (as indicated by decreasing
thickness of arrows in Figure~\ref{fig:vagueUse}).

@@

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
an example of the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, `accumulate' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of transmission biases are isolated, as in iterated learning models
(\citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
or when they interact with functional pressure towards efficient communication, for example as
formalised in the replicator mutator dynamic
(e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}).

The emphasis of previous models that studied the effects of transmission infidelity on the
evolution of language has been on inductive biases and the systematicity, compression and
regularisation that they can introduce. Here, we would like to show that transmission noise of
a different kind can lead to regularisation as well and also give rise to vague meaning. We
find that shared perceptual biases that perturb the transmission of successful signalling
behaviour can regularise, facilitate and accelerate the evolution of meaning conventions, albeit
at the cost of vagueness. Concretely, we formalise the expected change in the behaviour of a
population of agents that try to imitate other agents' signalling behaviour. We assume, however,
that both observation of others' behaviour, as well as realisation of behaviour, are both
systematically perturbed by noise. The resulting population-level dynamic generalises the
replicator dynamic \citepbjps{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
evolutionary dynamic based on imitation
(e.g.~\citealtbjps{Helbing1996:A-Stochastic-Be}; \citealtbjps{Schlag1998:Why-Imitate-and}). The inclusion of
confusability of stimuli does not undermine the possibility of evolving communicative signalling
behaviour. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
emergence of communicative signalling because it unifies and regularises evolutionary outcomes,
making it appear as if agents were applying inductive biases or generalising over partial
observations, when this is actually the sole effect of confusion of similar stimuli.

@@

The expected utility of choices at each choice point is:\footnote{The notation
  $\propto$, for `proportional to', that is used here and hereafter means that the right-hand
  side might still need to be normalised so as to have probabilities sum to one. In general,
  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\state' \in \States}
  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
    \States} P(\state \probbar \messg)  \myts
  U(\state, \state')\,, \text{\ \  where } &
P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
sum, for each interpretation $\state'$, of the probability that some random receiver will
choose $\state'$ given $\messg$ times the actual utility of pair $\state$ and
$\state'$.

The discrete-time replicator dynamic tracks changes in frequency of choices in the population
as proportional to their expected utilities
\citepbjps{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
  utilities are always positive.}
\begin{align}
  \label{eq:RDDiscrete}
  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto \Rstrat(\state \probbar \messg) \myts
    \EU(\state, \messg,\Sstrat)\,.   
\end{align}
For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
average agent in the population after the update is proportional to the probability of
seeing it before the update, which is $\Sstrat(\messg \probbar \state)$, times the
expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
by a gradient of current frequency and a measure of how good they are.

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioural dispositions in a population of signallers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioural adaptation. We focus
here on one of the simplest: imitation of success
\citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following: every now and then a random agent gets a chance to alter his behaviour for one
of his choice points (say this is a sender who gets to `reconsider' his choice of message
for state $\state$); the revising agent then observes what some random agent does at $\state$,
say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
expected utility of $\messg$ for state $\state$. See
Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
update scheme.

@@

Imitation of success, as described above, presupposes that agents make no mistakes when
observing states, or choosing interpretations. This may not always be an appropriate
assumption, especially when some states can be perceptually similar and therefore likely to be
confused for one another.
Human performance in these situations has been studied by a number of authors in experimental psychology. %~\citepbjps{takane_structures_1992}.
In a stimulus identification experiment~\citepbjps{luce-detection-1963}, subjects are presented in each trial with a stimulus to be identified out of a fixed set.
The more similar the stimuli are to each other, the larger the number of errors subjects make.
For example, in Robert Nosofsky's experiments~\citeyearparbjps{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the centre to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
In over 9000 trials, the two subjects identified the stimulus correctly only approximately 44\% and 35\% of the time.
Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colours, and magnitude of lines and areas~\citepbjpspre{see}{donkin-why-2015}.
The empirical data confirms not only the pervasiveness of variation in the subjects' ability to correctly identify stimuli, but also the relation between similarity and likelihood of stimulus confusion, and further suggests that the phenomenon might extend to all types of perception.

@@

The presence of observation and realisation noise also affects imitation of successes. Here, we
focus on the main ideas, formal detail is provided in Appendix~\ref{sec:imit-succ-with}. Take a
sender who gets to revise behaviour for state $\state$. What agents can plausibly revise by
imitation is their pure strategy, which maps perceived states onto messages. When an agent gets
to revise his strategy for the perceived state $\state$, this need not necessarily be the
actual state. Moreover, when that agent observes $\state$, another agent may perceive yet a
different state. Given what that latter agent perceives, his actual (pure) strategy will
determine what he plays. In sum, to describe how likely the potential imitator observes a
message choice $\messg$, we are interested in the conditional probability
$P_o(\messg \probbar \state)$ that some other random agent selects $\messg$ when the first
agent perceives $\state$. This $P_o(\messg \probbar \state)$ is derived from the prior
probability of states, the current sender population behaviour $\Sstrat$ and the given
observation noise (see Appendix~\ref{sec:imit-succ-with}). Eventually, the imitating agent adopts
$\messg$ as his choice for $\state$ with a probability given by the expected utility of sending
$\messg$ when perceiving state $\state$. Expected utility should, of course, take the
probabilistic confusability of states into account as well.
